Sampling Strategies in Decisions from Experience

### 2022-05-26 manuscript.Rmd
Final version: Discussion and Edits (SE)

\noindent
\fbox{
    \parbox{\textwidth}{
        \textbf{\hypertarget{box}{Box 1.}} 
        \textit{Reliance on Small Samples in DFE.}
        The binomial distribution of an outcome is the probability distribution of the number of times an outcome occurs in a sample.
        According to the \textit{de Moivre-Laplace theorem} (see, e.g., \protect\hyperlink{ref-georgiiStochastikEinfuhrungWahrscheinlichkeitstheorie2015}{Georgii, 2015}), the binomial distribution can be approximated by a symmetric normal distribution given a large sample.
        That is, the relative frequency with which an outcome occurs in a large sample is predicted to correspond to the objective probability of the outcome most of the time and is not predicted to be lower more often than higher or vice versa.
        However, for small samples, the binomial distribution associated with a small-probability outcome is positive skewed, causing the relative frequency with which this outcome occurs in the sample to be smaller rather than larger than its objective probability.
        That is, small-probability outcomes tend to be underrepresented in small samples.
        An explanation for the as-if underweighting of rare outcomes pattern may thus build on the assumption that the mind is sensitive to the information provided by small samples (\protect\hyperlink{ref-hertwigDecisionsExperienceEffect2004}{Hertwig et al., 2004}; \protect\hyperlink{ref-plonskyRelianceSmallSamples2015}{Plonsky et al., 2015}; see also \protect\hyperlink{ref-erevChoicePredictionCompetition2010}{Erev et al., 2010}, \protect\hyperlink{ref-erevAnomaliesForecastsDescriptive2017}{2017}, for the performance of models assuming reliance on small samples in prediction competitions).
    }
}
\newline
(ref:references-box-1) @georgiiStochastikEinfuhrungWahrscheinlichkeitstheorie2015; @hertwigDecisionsExperienceEffect2004; @erevChoicePredictionCompetition2010; @plonskyRelianceSmallSamples2015; @erevAnomaliesForecastsDescriptive2017

Yet, peoples' reliance on a small number of sampled outcomes in DFE, and the associated sampling error, are presumably not the sole cause of the description-experience gap [see @hauDescriptionexperienceGapRisky2008; @ungemachAreProbabilitiesOverweighted2009; @wulffMetaanalyticReviewTwo2018, who show that the as-if underweighting of rare outcomes pattern may appear, even if the overall number of sampled outcomes is large and the relative frequencies closely resemble the latent objective probabilities].
Rather, the specifics of the actual decision-making process---i.e, how the mind exercises the capacity of making decisions from experience as opposed to descriptions---might contribute to the gap.
In experimental studies of DFE using the *sampling paradigm*, people can, as much as they want, sequentially sample the outcomes of two latent prospects according to their objective probabilities before making a final, consequential choice between the prospects [see @hertwigDescriptionexperienceGapRisky2009, see also Fig.\ \@ref(fig:sampling-paradigm)].
Because in the sampling paradigm people cannot rely on the descriptions of outcomes and probabilities to make a decision, but are required to learn about the prospects from experiential sampling of outcomes over time, the tasks involved in the decision-making process differ from those in DFD.
That is, in order to model the process by which people arrive at the final choice in the sampling paradigm, one must explicate a number of assumptions concerning, in particular, the sampling process and the mind's operations on the sequence of sampled outcomes, which are of no relevance to DFD whatsoever, e.g.: How the mind

@@

\begin{figure}[t!]
\caption{Possible Sampling and Decision Strategies in the Sampling Paradigm and Their Impact on the Final Choice}
\includegraphics{manuscript_files/sampling-paradigm} \label{fig:sampling-paradigm}
\raggedright \textit{Note.} In the sampling paradigm, people are offered the choice between at least two prospects (here: violet vs. orange), whose exact properties (outcomes and probabilities) remain hidden. To learn about the properties, people can, as much as they want, sample the prospects' outcomes according to their probabilities. Sampling strategies: Refer to the sequential pattern with which people sample the outcomes from both prospects over time. People may switch infrequently (comprehensive sampling) or frequently (piecewise sampling) or something in between. Decision strategies: Refer to the way people integrate the sampled outcomes in order to evaluate the prospects. People may integrate all outcomes stemming from the same prospect in order to compute and compare summary scores (summary integration/decision); or they integrate and compare outcomes of the two prospects over repeated rounds (round-wise integration/decision).  
\end{figure}

@@

To summarize the preceding paragraphs: 
Since small samples may distort the information about the properties of latent prospects, a great deal of research on the sampling paradigm has been invested into the question which factors affect how much people sample [see @wulffAdaptiveExplorationWhat2019, Table 7.2, for a recent summary of relevant factors and papers].
Assuming that people proceed systematically when sampling outcomes, and that their sampling strategies are systematically related to decision strategies, Hills and Hertwig [-@hillsInformationSearchDecisions2010] emphasize that in order to understand how people make DFE and how this comes to imply certain choice patterns, one might need a richer understanding of peoples' sampling behavior.
That is, not only we need to know how much outcomes are sampled, but also in what way they are sampled, and how this affects the way the obtained information is processed to evaluate the latent prospects.

While offering a novel, convincing inroad into the broader question how sampling and information-processing can act together to shape DFE in environments where people can sample without costs,^[<!--Open footnote-->
Meaning that people do not gain/loose the amount of each sampled outcome. Of course, sampling implies other costs, such as time and cognitive capacities.
<!--Close footnote-->]
Hills and Hertwig's [-@hillsInformationSearchDecisions2010] work invites some objections, both theoretical and methodological.
The following gaps and problems stand out:

@@

3. *Use of heuristic analysis methods:*
The empirical re-analyses relies on various heuristic methods to assess whether the observed choices are consistent with the claimed links between sampling and decision strategies and their impact on the as-if weighting of rare events pattern.
For instance, aggregated data over participants and choice problems is used to compare, within and between sampling strategies, the proportions of choices that are consistent with the predictions of the decision strategies and an as-if underweighting of rare outcomes.
In addition, to obtain indicators for the applied sampling strategies, each participant's mean switching frequency across all choice problems was calculated and a median split was used to identify infrequently and frequently switching people, supposedly indicating comprehensive and piecewise sampling, respectively.
While acknowledging that existing data was used, such heuristic methods can be problematic, e.g., because they can cause a construct-behavior gap where inferences from the data to the tested theory, although statistically sound, become logically invalid.
In a rigorous analysis, @regenwetterConstructbehaviorGapBehavioral2017 demonstrate how the use of aggregate data and heuristic scoring rules, e.g., counting the proportion of correct choice predictions across problems and people, lead at least to hard-to-interpret inferences about individuals' as-if weighting patterns and the description-experience gap.
One can alleviate such risks by taking heterogeneity between individuals into account and modeling the as-if weighting more explicitly.
Also, the averaging of switching frequencies across choice problems to group into frequent and infrequent switchers can be falsely interpreted as if sampling strategies are more a characteristic of the decision maker than of the sampling process.
Finally, sampling error is only controlled for to the extent that choice trials in which small-probability outcomes were not sampled at least once are excluded from the analyses.
This still leaves room for the sampled relative frequencies to deviate from the latent objective probabilities.

@@

Because there is an uncountable number of decisions people confront in their daily life, behavioral decision research routinely abstracts from choices between particular alternatives, e.g., the choice between job offers, political parties to vote for, investment plans, and whatnot.
Rather, with the choice between at least two prospects, it studies a case which is counterfactual in that it omits the many particularities of each choice situation, but retains in the form of prospects the fundamental properties that almost all choice alternatives are assumed to possess.
These fundamental properties are considered the $n$ possible outcomes of a choice alternative, denoted as $x_1, ..., x_n$, and the probabilities with which these outcomes occur following the choice of the alternative, denoted as $p_X(x_1), ..., p_X(x_n)$. 
That is, each choice alternative that is fully described by a set of outcome-probability pairs, $\{(x_i, p_X(x_i))\}_{i \in \{1, ..., n\}}$, satisfying $\sum_{i = 1}^n p_X(x_i) = 1$, is a prospect.
It is only through these fundamental properties, or combinations thereof, that any necessary distinction between the prospects can be made.
Hence, systematic---as opposed to random---choices between prospects must be based on the assessment of differences in their properties.

Rarely in peoples' daily life, the information provided about the prospects take the form of explicit and complete descriptions of all their outcome-probability pairs.
In such rare cases, people would make DFD and can---strongly simplified---directly proceed by selecting a property or combinations thereof to assess differences among the prospects and make a respective choice.
In stark contrast and rather often, people are required to make DFE, where the mind can learn about the latent properties only by "experiential sampling [of outcomes] over time" [@hertwigDescriptionexperienceGapRisky2009, p. 517, square bracket added].
The sampling paradigm is just one paradigm that can be used to study DFE experimentally and allows people to sample the prospects' outcomes without costs and without constraints regarding the sequential pattern and the amount of sampled outcomes.

How do people arrive at a final choice in the sampling paradigm? 
Sequential sampling models of decision making under risk (DFD) and/or uncertainty [DFE, see @markantModelingChoiceSearch2015, who explicitly consider DFE and the sampling paradigm; see also, e.g., @bhatiaSequentialSamplingParadoxes2014; @busemeyerDecisionFieldTheory1993] assume that the decision-making process takes the form of an accumulation of evidence for the prospects over time.
The process is assumed to stop---i.e., a choice is made---once the accumulated evidence for one of the prospects reaches an evidence threshold.
More specifically, for DFE, it is assumed that the outcomes of prospects are sequentially sampled and integrated into dynamic decision variables and that a choice is made once the value of a decision variable exceeds a threshold.
How exactly this integration is carried out and what the value of a dynamic decision variable represents, is a theoretical question---and thus a matter of the exact specification of the model---largely concerned with the way the prospects are assumed to be evaluated.
Depending on the integration and evaluation mechanisms they assume, the member models of the sequential sampling framework have been shown to explain some of the robust deviations from EV maximization, including the as-if underweighting of rare outcomes in the sampling paradigm [@markantModelingChoiceSearch2015].
Moreover, these models implement speed-accuracy trade-offs, with the amount of evidence that must be sampled in favor for a prospect increasing with thresholds, leading to more accurate but slower choices for high thresholds, and vice versa.

@@

To implement the evidence accumulation process, the round-wise integration model assumes that each time after a *new* stochastic process was started and terminated on both prospects, their respective means are compared, with the prospect underlying the stochastic process with the greater mean receiving a round win.
Accordingly, assuming that $S^Y_m$ is the mean across $m$ outcomes successively sampled from prospect $P^Y$, this stage is modeled for the prospect $P^X$as the mapping
$$
\begin{aligned}
  Z^X : \mathbb{R} &\to \mathbb{N} \\
  S^X_t - S^Y_m &\mapsto 
  \begin{cases}
     1 &\text{if} \qquad  S^X_t > S^Y_m, \\
     0 &\text{else}
     \; ,
  \end{cases}
\end{aligned}
\tag{2}
$$
where $Z^X$ is a random variable that takes only the two values $0$ and $1$, denoting a lost or won mean comparison, respectively.
Let us denote the steps defined in (1) and (2) as a round of mean comparison (short: *comparison round*, see Fig.\ \@ref(fig:models)).

@@

The model assumes that an agent starts with equal probability to sample outcomes from one of the prospects and continues sampling while switching between prospects according to the switching probability $\psi$.
As in the round-wise integration model, the switching probability $\psi$ is assumed to be adopted prior the start of the sampling and accumulation process and is fixed throughout.

@@

Below, it is discussed which effects the key model parameters, i.e., the switching probability $\psi$ and the threshold $\theta$, could have on the choices and the corresponding as-if weighting patterns that are implied by the round-wise and the summary integration model.
The effects of the switching probability should be informative about the interplay between different sampling strategies and the integration/decision strategy each model resembles.
Assuming that both models aim to assess true differences in the latent properties of prospects, the effects of the threshold should indicate whether these differences are indeed more accurately assessed---thus, leading to more systematic choices and as-if weighting patterns---the more evidence is accumulated.
In addition, previous research has shown that differences as to how information about the properties of prospects is processed can translate to distinct signatures in CPT's psychoeconomic value and weighting function, without actually assuming that these functions are computed whatsoever [see @pachurHowTwainCan2017; @stewartDecisionSampling2006; @zilkerNonlinearProbabilityWeighting2021].^[<!--Open footnote-->
Note that it has been argued that the distinction between CPT's functions used to describe choice data and models aiming to explain how this data is generated in first place are reminiscent of Marr's [-@marrVisionComputationalInvestigation1982] different levels of explanation [@zilkerMeasuringModelingConstruction2020; also cf. @griffithsProbabilisticModelsCognition2010].
That is, whereas CPT may be considered a computational-level theory that describes an abstract solution to the problem the mind aims to solve when offered a choice between prospects, process models [cf. @jareckiFrameworkBuildingCognitive2020], such as heuristics and sequential sampling models, may be considered algorithmic-level theories that specify how the mind can approximate the abstract solution with the capacity limits it is constrained by.
<!--Close footnote-->]
Accordingly, this section also formulates expectations regarding the potential relations between switching probabilities and thresholds (data-generating parameters) and the CPT parameters (data-modeling parameters).

@@

Considering a *single* comparison round, the stochastic process for each prospect---which results in a mean computed across the outcomes sampled from the prospect---is predicted to terminate faster, the higher the probability that the subsequent outcome is sampled from the other prospect, i.e., the higher the switching probability $\psi$.
Because the switching probability is predicted to alter the length of the stochastic processes underlying *each independent* mean comparison, it should affect whether low-probability outcomes contribute to each required mean comparisons about as much as would be warranted based on their objective probabilities, or whether they contribute rather less.
That is, for high switching probabilities and thus, short stochastic processes, the binomial distribution associated with small-probability outcomes is positive skewed, causing the relative frequencies with which these outcomes occur within each comparison round to be smaller rather than larger than their latent objective probabilities.
Because such a distortion of the relative sampled frequencies from the objective probabilities is expected for each comparison round, increases in the threshold $\theta$ should not affect the following expectations concerning the model behavior; rather because each comparison is independent and increases in sample size (here: number of comparisons) reduce random error, the expectations should be more accurately met for higher thresholds:
If the small-probability outcome is smaller (larger) than the EV, high switching probabilities cause the mean within each comparison round to be an inflated (deflated) estimate of the EV.
Such an inflation and deflation of the means within each round can then translate to choices that take the form of an as-if underweighting of rare outcomes:
Specifically, if the means of the prospect with the lower (higher) EV are inflated (deflated), the outcomes of each mean comparison may be a reversal---in sign---of the outcome of a comparison between the EVs, eventually translating to choices resembling an as-if underweighting of rare outcomes.
Yet, if the means of the prospect with the lower (higher) EV are deflated (inflated), the difference between the means should be larger than the differences between the EV, which can simplify EV maximization [see @hertwigDecisionsExperienceWhy2010].
In contrast, for low values of the switching probability $\psi$---and therefore pairs of long stochastic processes---the means in each round should closely correspond to the EVs, leading to choices that are in line with EV maximization.

@@

Because the required (difference in) cumulative sums across all outcomes sampled from each prospect increases with thresholds $\theta$, the number of outcomes that is required to be sampled to reach the thresholds should increase accordingly.
Hence, in principle, the larger the thresholds, the more should small-probability outcomes contribute to the cumulative sum as much as would be warranted by their objective probabilities.
Accordingly, one might expect that the implied choices generally maximize the overall sampled mean and converge to the solutions of EV maximization, the larger the thresholds get.
However, the switching probabilities should introduce temporal imbalances between prospects regarding the number of sampled outcomes over which the sums are computed (and compared, assuming relative thresholds, respectively).
That is, whatever prospect is first sampled from is allowed to start accumulating evidence sooner than the respective other prospect.
This is likely to be a negligible factor for large but not for small switching probabilities.
Specifically, the temporal imbalances---i.e., the temporal advantage for the prospect which is first sampled from---grows as the switching probabilities decrease.
Accordingly, for small switching probabilities, choices might be largely driven by the truly random event which prospect is first sampled from.
This random choice behavior should diminish with increases in the switching probability and choices should converge to the solutions of sampled mean and EV maximization the higher the thresholds get.

@@

Before elaborating how the expectations outlined above may translate into estimates of the CPT parameters, the following section reviews how CPT in general and its weighting function in particular capture deviations from EV maximization.

@@

A test set of 60 choice problems was obtained by stratified sampling from an initial set of 10,000 problems.
The stratification procedure was used to ensure that the test set contained choice problems with small-probability outcomes of different rank---i.e., small-probability outcomes that are either higher or lower than the EV of the risky prospect---as well as choice problems without small-probability outcomes.
The exact procedure was as follows:
For each of the 10,000 problems in the initial set, three outcomes were randomly drawn from a uniform distribution over the interval $[0, 20]$, with the smallest and highest of these three outcomes being assigned to the risky prospect to omit dominance.^[I.e., when one prospect always produces a higher outcome.]
The probability for the lower outcome of the risky prospect, $p_x(x_{low})$, was drawn from a uniform distribution over the interval $(0, 1)$, with the probability of the higher outcome being set to $p_x(x_{high}) = 1-p_x(x_{low})$.
To obtain the 60 choice problems of the test set, the initial set was divided into three subsets, where each subset contained either all problems with $p_x(x_{high}) \in (0,.2)$ or $p_x(x_{high}) \in [.2,.8]$ or $p_x(x_{high}) \in (.8,1)$.
From each of these subset, then, $20$ choice problems were randomly sampled.

@@

Figure\ \@ref(fig:sample-sizes) displays the number of outcomes sampled over the course of a sampling and accumulation process for each combination of switching probabilities and thresholds.
The model predicts a decrease in the overall number of sampled outcomes with increases in the switching probability, whereas the overall number of sampled outcomes is predicted to increase with thresholds at relatively constant rate.
These differences in the number of sampled outcomes are subtle indicators for how sampling strategies and the mechanisms of the round-wise integration model act together to imply different final choices.
Essentially, they foreshadow how the round-wise integration model accumulates evidence for differences in the latent EV for low switching probabilities, and for differences in the latent probability to produce a higher outcome most of the time for high switching probabilities.

@@

# Discussion

@@

>>>>

\noindent
\fbox{
    \parbox{\textwidth}{
        \textbf{\hypertarget{box}{Box 1.}} 
        \textit{Reliance on Small Samples in DFE.}
        The binomial distribution of an outcome is the probability distribution of the number of times an outcome occurs in a sample.
        According to the \textit{de Moivre-Laplace theorem} (see, e.g., \protect\hyperlink{ref-georgiiStochastikEinfuhrungWahrscheinlichkeitstheorie2015}{Georgii, 2015}), the binomial distribution can be approximated by a symmetric normal distribution given a large sample.
        That is, the relative frequency with which an outcome occurs in a large sample is predicted to correspond to the objective probability of the outcome most of the time and is not predicted to be lower more often than higher or vice versa.
        However, for small samples, the binomial distribution associated with a small-probability outcome is positively skewed, causing the relative frequency with which this outcome occurs in the sample to be smaller rather than larger than its objective probability.
        That is, small-probability outcomes tend to be underrepresented in small samples.
        An explanation for the as-if underweighting of rare outcomes pattern may thus build on the assumption that the mind is sensitive to the information provided by small samples (\protect\hyperlink{ref-hertwigDecisionsExperienceEffect2004}{Hertwig et al., 2004}; \protect\hyperlink{ref-plonskyRelianceSmallSamples2015}{Plonsky et al., 2015}; see also \protect\hyperlink{ref-erevChoicePredictionCompetition2010}{Erev et al., 2010}, \protect\hyperlink{ref-erevAnomaliesForecastsDescriptive2017}{2017}, for the performance of models assuming reliance on small samples in prediction competitions).
    }
}
\newline
(ref:references-box-1) @georgiiStochastikEinfuhrungWahrscheinlichkeitstheorie2015; @hertwigDecisionsExperienceEffect2004; @erevChoicePredictionCompetition2010; @plonskyRelianceSmallSamples2015; @erevAnomaliesForecastsDescriptive2017

Yet, peoples' reliance on a small number of sampled outcomes in DFE, and the associated sampling error are presumably not the sole cause of the description-experience gap [see @hauDescriptionexperienceGapRisky2008; @ungemachAreProbabilitiesOverweighted2009; @wulffMetaanalyticReviewTwo2018, who show that the as-if underweighting of rare outcomes pattern may appear, even if the overall number of sampled outcomes is large and the relative frequencies closely resemble the latent objective probabilities].
Rather, the specifics of the actual decision-making process---i.e, how the mind exercises the capacity of making decisions from experience as opposed to descriptions---might contribute to the gap.
In experimental studies of DFE using the *sampling paradigm*, people can, as much as they want, sequentially sample the outcomes of two latent prospects according to their objective probabilities before making a final, consequential choice between the prospects [see @hertwigDescriptionexperienceGapRisky2009, see also Fig.\ \@ref(fig:sampling-paradigm)].
Because in the sampling paradigm people cannot rely on the descriptions of outcomes and probabilities to make a decision, but are required to learn about the prospects from experiential sampling of outcomes over time, the tasks involved in the decision-making process differ from those in DFD.
That is, in order to model the process by which people arrive at the final choice in the sampling paradigm, one must explicate a number of assumptions concerning, in particular, the sampling process and the mind's operations on the sequence of sampled outcomes, which are of no relevance to DFD whatsoever, e.g.: How the mind

@@

\begin{figure}[t!]
\caption{Possible Sampling and Decision Strategies in the Sampling Paradigm and Their Impact on the Final Choice}
\includegraphics{manuscript_files/sampling-paradigm} \label{fig:sampling-paradigm}
\raggedright \textit{Note.} In the sampling paradigm, people are offered the choice between at least two prospects (here: violet vs. orange), whose exact properties (outcomes and probabilities) remain hidden. To learn about the properties, people can, as much as they want, sample the prospects' outcomes according to their probabilities. Sampling strategies: Refer to the sequential pattern with which people sample the outcomes from both prospects over time. People may switch infrequently (comprehensive sampling) or frequently (piecewise sampling) or with a frequency somewhere in between. Decision strategies: Refer to the way people integrate the sampled outcomes in order to evaluate the prospects. People may integrate all outcomes stemming from the same prospect in order to compute and compare summary scores (summary integration/decision); or they integrate and compare outcomes of the two prospects over repeated rounds (round-wise integration/decision).  
\end{figure}

@@

To summarize the preceding paragraphs: 
Since small samples may distort the information about the properties of latent prospects, a great deal of research on the sampling paradigm has been invested into the question of which factors affect how much people sample [see @wulffAdaptiveExplorationWhat2019, Table 7.2, for a recent summary of relevant factors and papers].
Assuming that people proceed systematically when sampling outcomes, and that their sampling strategies are systematically related to decision strategies, Hills and Hertwig [-@hillsInformationSearchDecisions2010] emphasize that in order to understand how people make DFE and how this comes to imply certain choice patterns, one might need a richer understanding of peoples' sampling behavior.
That is, not only do we need to know how many outcomes are sampled, but also in what way they are sampled, and how this affects the way the obtained information is processed to evaluate the latent prospects.

While offering a novel, convincing inroad into the broader question of how sampling and information-processing can act together to shape DFE in environments where people can sample without costs,^[<!--Open footnote-->
Meaning that people do not gain/loose the amount of each sampled outcome. Of course, sampling implies other costs, such as time and cognitive capacities.
<!--Close footnote-->]
Hills and Hertwig's [-@hillsInformationSearchDecisions2010] work invites some objections, both theoretical and methodological.
The following gaps and problems stand out:

@@

3. *Use of heuristic analysis methods:*
The empirical re-analysis relies on various heuristic methods to assess whether the observed choices are consistent with the claimed links between sampling and decision strategies and their impact on the as-if weighting of rare events pattern.
For instance, aggregated data over participants and choice problems is used to compare, within and between sampling strategies, the proportions of choices that are consistent with the predictions of the decision strategies and an as-if underweighting of rare outcomes.
In addition, to obtain indicators for the applied sampling strategies, each participant's mean switching frequency across all choice problems was calculated and a median split was used to identify infrequently and frequently switching people, supposedly indicating comprehensive and piecewise sampling, respectively.
While acknowledging that existing data was used, such heuristic methods can be problematic, e.g., because they can cause a construct-behavior gap where inferences from the data to the tested theory, although statistically sound, become logically invalid.
In a rigorous analysis, @regenwetterConstructbehaviorGapBehavioral2017 demonstrate how the use of aggregated data and heuristic scoring rules, e.g., counting the proportion of correct choice predictions across problems and people, lead at least to hard-to-interpret inferences about individuals' as-if weighting patterns and the description-experience gap.
One can alleviate such risks by taking heterogeneity between individuals into account and modeling the as-if weighting more explicitly.
Also, the averaging of switching frequencies across choice problems to group into frequent and infrequent switchers can be falsely interpreted as if sampling strategies are more a characteristic of the decision maker than of the sampling process.
Finally, sampling error is only controlled for to the extent that choice trials in which small-probability outcomes were not sampled at least once are excluded from the analyses.
This still leaves room for the sampled relative frequencies to deviate from the latent objective probabilities.

@@

Because there is an uncountable number of decisions people confront in their daily life, behavioral decision research routinely abstracts from choices between particular alternatives, e.g., the choice between job offers, political parties to vote for, or investment plans.
Rather, with the choice between at least two prospects, it studies a case which is counterfactual in that it omits the many particularities of each choice situation but retains in the form of prospects the fundamental properties that almost all choice alternatives are assumed to possess.
These fundamental properties are considered the $n$ possible outcomes of a choice alternative, denoted as $x_1, ..., x_n$, and the probabilities with which these outcomes occur following the choice of the alternative, denoted as $p_X(x_1), ..., p_X(x_n)$. 
That is, each choice alternative that is fully described by a set of outcome-probability pairs, $\{(x_i, p_X(x_i))\}_{i \in \{1, ..., n\}}$, satisfying $\sum_{i = 1}^n p_X(x_i) = 1$, is a prospect.
It is only through these fundamental properties, or combinations thereof, that any necessary distinction between the prospects can be made.
Hence, systematic---as opposed to random---choices between prospects must be based on the assessment of differences in their properties.

Rarely in peoples' daily life, does the information provided about the prospects take the form of explicit and complete descriptions of all their outcome-probability pairs.
In such rare cases, people would make DFD and can---strongly simplified---directly proceed by selecting a property or combinations thereof to assess differences among the prospects and make a respective choice.
In stark contrast and rather often, people are required to make DFE, where the mind can learn about the latent properties only by "experiential sampling [of outcomes] over time" [@hertwigDescriptionexperienceGapRisky2009, p. 517, square bracket added].
The sampling paradigm is just one paradigm that can be used to study DFE experimentally and allows people to sample the prospects' outcomes without costs and without constraints regarding the sequential pattern and the amount of sampled outcomes.

How do people arrive at a final choice in the sampling paradigm? 
Sequential sampling models of decision making under risk (DFD) and/or uncertainty [DFE, see @markantModelingChoiceSearch2015, who explicitly consider DFE and the sampling paradigm; see also, e.g., @bhatiaSequentialSamplingParadoxes2014; @busemeyerDecisionFieldTheory1993] assume that the decision-making process takes the form of an accumulation of evidence for the prospects over time.
The process is assumed to stop---i.e., a choice is made---once the accumulated evidence for one of the prospects reaches an evidence threshold.
More specifically, for DFE, it is assumed that the outcomes of prospects are sequentially sampled and integrated into dynamic decision variables and that a choice is made once the value of a decision variable exceeds a threshold.
How exactly this integration is carried out and what the value of a dynamic decision variable represents, is a theoretical question---and thus a matter of the exact specification of the model---largely concerned with the way the prospects are assumed to be evaluated.
Depending on the integration and evaluation mechanisms they assume, the member models of the sequential sampling framework have been shown to explain some of the robust deviations from EV maximization, including the as-if underweighting of rare outcomes in the sampling paradigm [@markantModelingChoiceSearch2015].
Moreover, these models implement speed-accuracy trade-offs, with the amount of evidence that must be sampled in favor of a prospect increasing with thresholds, leading to more accurate but slower choices for high thresholds, and vice versa.

@@

To implement the evidence accumulation process, the round-wise integration model assumes that each time a *new* stochastic process gets started and then terminated on both prospects, their respective means are compared, with the prospect underlying the stochastic process with the greater mean receiving a round win.
Accordingly, assuming that $S^Y_m$ is the mean across $m$ outcomes successively sampled from prospect $P^Y$, this stage is modeled for the prospect $P^X$as the mapping
$$
\begin{aligned}
  Z^X : \mathbb{R} &\to \mathbb{N} \\
  S^X_t - S^Y_m &\mapsto 
  \begin{cases}
     1 &\text{if} \qquad  S^X_t > S^Y_m, \\
     0 &\text{else}
     \; ,
  \end{cases}
\end{aligned}
\tag{2}
$$
where $Z^X$ is a random variable that takes only the two values $0$ and $1$, denoting a lost or won mean comparison, respectively.
Let us denote the steps defined in (1) and (2) as a round of mean comparison (short: *comparison round*, see Fig.\ \@ref(fig:models)).

@@

The model assumes that an agent starts with equal probability to sample outcomes from one of the prospects and continues sampling while switching between prospects according to the switching probability $\psi$.
As in the round-wise integration model, the switching probability $\psi$ is assumed to be adopted prior to the start of the sampling and accumulation process and is fixed throughout.

@@

Below, it is discussed which effects the key model parameters, i.e., the switching probability $\psi$ and the threshold $\theta$, could have on the choices and the corresponding as-if weighting patterns that are implied by the round-wise and the summary integration model.
The effects of the switching probability should be informative about the interplay between different sampling strategies and the integration/decision strategy each model resembles.
Assuming that both models aim to assess true differences in the latent properties of prospects, the effects of the threshold should indicate whether these differences are indeed more accurately assessed---thus, leading to more systematic choices and as-if weighting patterns---the more evidence is accumulated.
In addition, previous research has shown that differences as to how information about the properties of prospects is processed can translate to distinct signatures in CPT's psychoeconomic value and weighting function, without actually assuming that these functions are computed whatsoever [see @pachurHowTwainCan2017; @stewartDecisionSampling2006; @zilkerNonlinearProbabilityWeighting2021].^[<!--Open footnote-->
Note that it has been argued that the distinction between CPT's functions used to describe choice data and models aiming to explain how this data is generated in the first place are reminiscent of Marr's [-@marrVisionComputationalInvestigation1982] different levels of explanation [@zilkerMeasuringModelingConstruction2020; also cf. @griffithsProbabilisticModelsCognition2010].
That is, whereas CPT may be considered a computational-level theory that describes an abstract solution to the problem the mind aims to solve when offered a choice between prospects, process models [cf. @jareckiFrameworkBuildingCognitive2020], such as heuristics and sequential sampling models, may be considered algorithmic-level theories that specify how the mind can approximate the abstract solution with the capacity limits it is constrained by.
<!--Close footnote-->]
Accordingly, this section also formulates expectations regarding the potential relations between switching probabilities and thresholds (data-generating parameters) and the CPT parameters (data-modeling parameters).

@@

Considering a *single* comparison round, the stochastic process for each prospect---which results in a mean computed across the outcomes sampled from the prospect---is predicted to terminate faster, the higher the probability that the subsequent outcome is sampled from the other prospect, i.e., the higher the switching probability $\psi$.
Because the switching probability is predicted to alter the length of the stochastic processes underlying *each independent* mean comparison, it should affect whether low-probability outcomes contribute to each required mean comparison about as much as would be warranted based on their objective probabilities, or whether they contribute rather less.
That is, for high switching probabilities and thus, short stochastic processes, the binomial distribution associated with small-probability outcomes is positively skewed, causing the relative frequencies with which these outcomes occur within each comparison round to be smaller rather than larger than their latent objective probabilities.
Because such a distortion of the relative sampled frequencies from the objective probabilities is expected for each comparison round, increases in the threshold $\theta$ should not affect the following expectations concerning the model behavior; rather because each comparison is independent and increases in sample size (here: number of comparisons) reduce random error, the expectations should be more accurately met for higher thresholds:
If the small-probability outcome is smaller (larger) than the EV, high switching probabilities cause the mean within each comparison round to be an inflated (deflated) estimate of the EV.
Such an inflation and deflation of the means within each round can then translate to choices that take the form of an as-if underweighting of rare outcomes:
Specifically, if the means of the prospect with the lower (higher) EV are inflated (deflated), the outcomes of each mean comparison may be a reversal---in sign---of the outcome of a comparison between the EVs, eventually translating to choices resembling an as-if underweighting of rare outcomes.
Yet, if the means of the prospect with the lower (higher) EV are deflated (inflated), the difference between the means should be larger than the differences between the EV, which can simplify EV maximization [see @hertwigDecisionsExperienceWhy2010].
In contrast, for low values of the switching probability $\psi$---and therefore pairs of long stochastic processes---the means in each round should closely correspond to the EVs, leading to choices that are in line with EV maximization.

@@

Because the required (difference in) cumulative sums across all outcomes sampled from each prospect increases with thresholds $\theta$, the number of outcomes that is required to be sampled to reach the thresholds should increase accordingly.
Hence, in principle, the larger the thresholds, the more should small-probability outcomes contribute to the cumulative sum as much as would be warranted by their objective probabilities.
Accordingly, one might expect that the implied choices generally maximize the overall sampled mean and converge to the solutions of EV maximization, the larger the thresholds get.
However, the switching probabilities should introduce temporal imbalances between prospects regarding the number of sampled outcomes over which the sums are computed (and compared, assuming relative thresholds, respectively).
That is, whatever prospect is first sampled from is allowed to start accumulating evidence sooner than the respective other prospect.
This is likely to be a negligible factor for large but not for small switching probabilities.
Specifically, the temporal imbalances---i.e., the temporal advantage for the prospect which is sampled from first---grows as the switching probabilities decrease.
Accordingly, for small switching probabilities, choices might be largely driven by the truly random event which prospect is first sampled from.
This random choice behavior should diminish with increases in the switching probability and choices should converge to the solutions of sampled mean and EV maximization the higher the thresholds get.

@@

Before elaborating on how the expectations outlined above may translate into estimates of the CPT parameters, the following section reviews how CPT in general and its weighting function in particular capture deviations from EV maximization.

@@

A test set of 60 choice problems was obtained by stratified sampling from an initial set of 10,000 problems.
The stratification procedure was used to ensure that the test set contained choice problems with small-probability outcomes of different rank---i.e., small-probability outcomes that are either higher or lower than the EV of the risky prospect---as well as choice problems without small-probability outcomes.
The exact procedure was as follows:
For each of the 10,000 problems in the initial set, three outcomes were randomly drawn from a uniform distribution over the interval $[0, 20]$, with the smallest and highest of these three outcomes being assigned to the risky prospect to omit dominance.^[<!--Open footnote-->I.e., where one prospect always produces a higher outcome.<!--Close footnote-->]
The probability for the lower outcome of the risky prospect, $p_x(x_{low})$, was drawn from a uniform distribution over the interval $(0, 1)$, with the probability of the higher outcome being set to $p_x(x_{high}) = 1-p_x(x_{low})$.
To obtain the 60 choice problems of the test set, the initial set was divided into three subsets, where each subset contained either all problems with $p_x(x_{high}) \in (0,.2)$ or $p_x(x_{high}) \in [.2,.8]$ or $p_x(x_{high}) \in (.8,1)$.
From each of these subsets, then, $20$ choice problems were randomly sampled.

@@

Figure\ \@ref(fig:sample-sizes) displays the number of outcomes sampled over the course of a sampling and accumulation process for each combination of switching probabilities and thresholds.
The model predicts a decrease in the overall number of sampled outcomes with increases in the switching probability, whereas the overall number of sampled outcomes is predicted to increase with thresholds at a relatively constant rate.
These differences in the number of sampled outcomes are subtle indicators for how sampling strategies and the mechanisms of the round-wise integration model act together to imply different final choices.
Essentially, they foreshadow how the round-wise integration model accumulates evidence for differences in the latent EV for low switching probabilities, and for differences in the latent probability to produce a higher outcome most of the time for high switching probabilities.

@@

# Discussion and Conclusion

@@

\noindent
I thank Thorsten Pachur and Veronika Zilker for the opportunity of an internship at the Center for Adaptive Rationality at the Max Planck Institute for Human Development in Berlin and for supervising this thesis.
